---
title: "Industry Opportunity and Workforce Equity"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

This report identifies industries with projected employment growth and median wages above the living-wage threshold, then compares their workforce composition with the overall workforce age 16 and older.

```{r setup}
# Quarto evaluates this document from its containing Industry directory.
# The sourced analysis script uses paths relative to that same directory.
source("industry_analysis.R")
source("industry_vizes.R")

library(data.table)
library(dplyr)
library(ggplot2)
```

<!-- ## Industry concentration crosstab -->

<!-- Each `*_diff` field is the industry's share for that race/ethnicity group minus the overall workforce share. Positive values indicate overrepresentation. Each `*_z` field is the corresponding approximate Z-score; larger absolute values indicate stronger evidence that the difference is distinct from zero. -->

<!-- ```{r concentration-crosstab} -->

<!-- #| tbl-cap: "Race/ethnicity concentration differences and Z-scores by industry" -->

<!-- crosstab_display <- copy(naics_conc_crosstab) -->

<!-- setnames(crosstab_display, "sector", "Industry") -->

<!-- numeric_cols <- names(crosstab_display)[vapply(crosstab_display, is.numeric, logical(1))] -->

<!-- crosstab_display[, (numeric_cols) := lapply(.SD, round, digits = 3), .SDcols = numeric_cols] -->

<!-- knitr::kable(crosstab_display, format = "html", digits = 3) -->

<!-- ``` -->

## Workforce equity overview

Bubble position shows the difference between each industry's POC and female shares and the overall workforce shares. Bubble area represents projected 2034 employment; fill represents median annual wage.

```{r equity-bubble, fig.width=12, fig.height=8}
#| fig-cap: "POC and female workforce concentration by industry"

make_bubble_plot_naics(filter(naics_stats, cagr_24_34 > 0.01))
```

## Female and POC share differences

The paired columns compare each industry's female and POC shares with the corresponding shares in the overall workforce. Industries follow the same display order as the table and heatmap.

```{r equity-dot-display, fig.width=12, fig.height=10}
#| fig-cap: "Female and POC workforce share differences by industry"

make_equity_dot_plot_naics(filter(naics_stats, cagr_24_34 > 0.01))
```

## Race and ethnicity concentration heatmap

This view shows the direction and size of each industry's concentration difference across all race/ethnicity groups.

```{r concentration-heatmap, fig.width=12, fig.height=9}
#| fig-cap: "Race/ethnicity share differences relative to the overall workforce"

heatmap_data <- as.data.frame(filter(naics_conc_long, cagr_24_34 > 0.01)) %>%
  mutate(
    sector = factor(sector, levels = rev(unique(naics_stats$sector))),
    prace_adj = factor(prace_adj)
  )

ggplot(heatmap_data, aes(x = prace_adj, y = sector, fill = diff_share)) +
  geom_tile(color = "white", linewidth = 0.25) +
  geom_text(aes(label = scales::percent(diff_share, accuracy = 0.1)), size = 3) +
  scale_fill_gradient2(
    low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0,
    labels = scales::label_percent(accuracy = 1),
    name = "Share difference"
  ) +
  labs(
    x = NULL,
    y = NULL,
    subtitle = "Positive values indicate overrepresentation in the industry."
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "bottom"
  )
```