---
title: "Industry Opportunity and Workforce Equity"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

This report identifies industries with projected employment growth and median wages above the living-wage threshold, then compares their workforce composition with the overall workforce age 16 and older.

```{r setup}
# Quarto evaluates this document from its containing Industry directory.
# The sourced analysis script uses paths relative to that same directory.
source("industry_analysis.R")
source("industry_vizes.R")
source("../render_table.R")

library(data.table)
library(dplyr)
library(ggplot2)
```

## Workforce equity overview

Bubble position shows the difference between each industry's POC and female shares and the overall workforce shares. Bubble area represents projected 2034 employment; fill represents median annual wage.

```{r equity-bubble, fig.width=12, fig.height=8}
#| fig-cap: "POC and female workforce concentration by industry"

make_bubble_plot_naics(filter(naics_stats, cagr_24_34 > 0.01))
```

## Female and POC share differences

The paired columns compare each industry's female and POC shares with the corresponding shares in the overall workforce. The right-hand bars show median annual wages on a shared scale, with pale yellow indicating lower wages and midnight blue indicating higher wages. Industries follow the same display order as the table and heatmap.

```{r equity-dot-display, fig.width=12, fig.height=10}
#| fig-cap: "Female and POC workforce share differences and median annual wages by industry"

make_equity_dot_plot_naics(filter(naics_stats, cagr_24_34 > 0.01))
```

## Race and ethnicity concentration heatmap

This view shows the direction and size of each industry's concentration difference across all race/ethnicity groups.

```{r concentration-heatmap, fig.width=12, fig.height=9}
#| fig-cap: "Race/ethnicity share differences relative to the overall workforce"

heatmap_data <- as.data.frame(filter(naics_conc_long, cagr_24_34 > 0.01)) %>%
  mutate(
    sector = factor(sector, levels = rev(unique(naics_stats$sector))),
    prace_adj = factor(prace_adj)
  )

ggplot(heatmap_data, aes(x = prace_adj, y = sector, fill = diff_share)) +
  geom_tile(color = "white", linewidth = 0.25) +
  geom_text(aes(label = scales::percent(diff_share, accuracy = 0.1)), size = 3) +
  scale_fill_gradient2(
    low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0,
    labels = scales::label_percent(accuracy = 1),
    name = "Share difference"
  ) +
  labs(
    x = NULL,
    y = NULL,
    subtitle = "Positive values indicate overrepresentation in the industry."
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "bottom"
  )
```

## Industry concentration crosstab

Each `*_diff` field is the industry's share for that race/ethnicity group minus the overall workforce share. Positive values indicate overrepresentation. Each `*_z` field is the corresponding approximate Z-score; larger absolute values indicate stronger evidence that the difference is distinct from zero.


## Sector summary

This table includes all sectors in `naics_stats`, including those below the growth threshold used in the visualizations. Annual growth is the compound annual growth rate for 2024–2034.

```{r sector-summary, echo=FALSE}
render_data_table(
  naics_stats[emp_2034 > 5000, .(
    Sector = sector,
    `Employment, 2024` = emp_2024,
    `Employment, 2034` = emp_2034,
    `Annual growth (%)` = round(100 * cagr_24_34, 1),
    `Median annual wage ($)` = round(WAGP_median)
  )],
  caption = "Sector summary",
  digits = 1
)
```
