---
title: "High-growth Occupation Opportunity and Workforce Equity"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

This report identifies 3-digit SOC occupation groups projected to grow at annual rates over 1 percent through 2034, then compares all workers in each group with the overall workforce age 16 and older. Employment and projected openings are summed across detailed occupations; growth and survey-weighted median wages are calculated at the group level.

```{r setup}
# Quarto evaluates this document from its containing Occupation directory.
filter_type <- "high_growth"
source("occupation_analysis.R")
source("occupation_vizes.R")
source("../render_table.R")

library(data.table)
library(dplyr)
library(ggplot2)
```

## Workforce equity overview

Bubble position shows the difference between each occupation group's POC and female shares and the overall workforce shares. Bubble area represents projected openings through 2034; fill represents median annual wage.

```{r equity-bubble, fig.width=12, fig.height=8}
#| fig-cap: "POC and female workforce concentration by occupation"

make_focus_bubble_plot(soc_stats)
```

## Female and POC share differences

The paired columns compare each SOC3 occupation group's female and POC shares with the corresponding shares in the overall workforce.

```{r equity-dot-display, fig.width=12, fig.height=max(8, 0.4*nrow(soc_stats))}
#| fig-cap: "Female and POC workforce share differences by SOC3 occupation group"

soc3_equity_stats <- copy(soc_stats) %>% setorder(soc3_code)

make_equity_dot_plot_soc3(soc3_equity_stats)
```

## Race and ethnicity concentration heatmap

This view shows the direction and strength of each occupation group's concentration difference across all race/ethnicity groups.

```{r concentration-heatmap, fig.width=12, fig.height=max(8, 0.4*nrow(soc_stats))}
#| fig-cap: "Race/ethnicity concentration relative to the overall workforce"

make_race_conc_heatmap_z(race_conc_long, soc_stats, openings_min = 0)
```

## Occupation-group summary

Projected openings are totals for 2024–2034. The education percentage is weighted by projected openings; unknown requirements remain in the denominator and do not count as requiring less than a bachelor's degree.

```{r group-summary, echo=FALSE}
render_data_table(
  soc_stats[, .(
    Occupation = Label,
    `SOC3` = soc3_code,
    `Employment, 2024` = emp_2024,
    `Employment, 2034` = emp_2034,
    `Annual growth (%)` = round(100 * cagr_24_34, 1),
    `Openings, 2024–2034` = openings_total_24_34,
    `Median annual wage ($)` = round(WAGP_median),
    `Openings below bachelor's (%)` = round(100 * below_bachelors_share, 1),
    `Education unknown (%)` = round(100 * education_unknown_share, 1)
  )],
  caption = "Occupation group summary",
  digits = 1
)
```
