---
title: "Annual Trend Data W/O Race & Ethnicity Details"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
library(dplyr)
library(data.table)

save_data <- TRUE

if(save_data) 
    library(writexl)

# Note: one has to set the data path in get_pums_res_equity.R
lapply(c("get_pums_res_equity.R",
         "plot_rates.R", "plot_counts.R"), source)

dyears <- c(2010:2019, 2021:2024)

pumslist_h <- list()
pumslist_h <- lapply(dyears, get_pums_res_equity_h, span = 1)
pumslist_h <- lapply(pumslist_h, prep_internet_data) %>%
              lapply(prep_renter_burden_data)

pumslist_p <- list()
pumslist_p <- lapply(dyears, get_pums_res_equity_p, span = 1)
pumslist_p <- lapply(pumslist_p, prep_wage_data) %>%
              lapply(prep_labor_force_data) %>%
              lapply(prep_health_insurance_data) %>%
              lapply(prep_poverty_data)

results <- list() # for collecting all results to store as spreadsheet
```

## Population Indicators

### Population

```{r}
#| label: population_data
rs_pop <- lapply(pumslist_p, psrc_pums_count) %>% data.table::rbindlist(fill=TRUE)
rs_pop[, `:=`(share = NULL, share_moe = NULL)]
results[["population"]] <- rs_pop
```



```{r}
#| label: population_plot
#p_pop <- plot_counts(rs_pop, "DATA_YEAR", "count", moe = "count_moe")
p_pop <- plot_line_counts(rs_pop, "DATA_YEAR", "count", 
                          legend = "region's population")
plot(p_pop)
```

### Median household income

```{r}
#| label: income_data
rs_hhincome <- lapply(pumslist_h, psrc_pums_median, stat_var = "HINCP") %>% 
    data.table::rbindlist(fill=TRUE)
results[["income"]] <- rs_hhincome
```


```{r}
#| label: income_plot
#p_hhincome <- plot_counts(rs_hhincome, "DATA_YEAR", "HINCP_median", moe = "HINCP_median_moe")
p_hhincome <- plot_line_counts(rs_hhincome, "DATA_YEAR", "HINCP_median", legend = "region's median income")
plot(p_hhincome)
```

## Poverty Indicators

### Poverty rates

```{r}
#| label: poverty_data
rs_poverty <- lapply(pumslist_p, psrc_pums_count, group_vars = "poverty", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
results[["poverty"]] <- rs_poverty[poverty == "Yes"]
```


```{r}
#| label: poverty_plot
#p_poverty <- plot_rates(filter(rs_poverty, poverty == "Yes"),
#    group_vars = "DATA_YEAR", whiskers = TRUE) # error bars 
p_poverty <- plot_line_rates(
    results[["poverty"]],
    xcol = "DATA_YEAR",
    legend = "region's poverty rates"
)

plot(p_poverty)
```


### Low income rates

```{r}
#| label: low_income_data
rs_lowinc <- lapply(pumslist_p, psrc_pums_count, group_vars = "low_income", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
results[["low_income"]] <- rs_lowinc[low_income == "Yes"]
```

```{r}
#| label: low_income_plot
#p_lowinc <- plot_rates(
#    filter(rs_lowinc, low_income == "Yes"),
#    group_vars = "DATA_YEAR",
#    whiskers = TRUE # error bars 
#)

p_lowinc <- plot_line_rates(
    results[["low_income"]],
    xcol = "DATA_YEAR",
    legend = "region's low income rates"
)

plot(p_lowinc)
```

## Employment

### Labor force participation (16 and over)

```{r}
#| label: labor_force_data
rs_lfp <- lapply(pumslist_p, psrc_pums_count, group_vars = "labor_force_status", incl_na=FALSE) %>%
   data.table::rbindlist(fill=TRUE)
results[["labor_force"]] <- rs_lfp[labor_force_status == "In labor force"]
```

```{r}
#| label: labor_force_plot
#p_lfp <- plot_rates(
#    filter(rs_lfp, labor_force_status == "In labor force"),
#    group_vars = "DATA_YEAR",
#    whiskers = TRUE # error bars 
#)

p_lfp <- plot_line_rates(
    results[["labor_force"]],
    xcol = "DATA_YEAR",
    legend = "region's in-labor force"
)

plot(p_lfp)
```

### Unemployment

```{r}
#| label: unemployment_data
rs_unemp <- lapply(pumslist_p, psrc_pums_count, group_vars = "employment_status", incl_na=FALSE) %>%
   data.table::rbindlist(fill=TRUE)
results[["unemployment"]] <- rs_unemp[employment_status == "Unemployed"]
```

```{r}
#| label: unemployment_plot
#p_unemp <- plot_rates(
#    filter(rs_unemp, employment_status == "Unemployed"),
#    group_vars = "DATA_YEAR",
#    whiskers = TRUE # error bars 
#)

p_unemp <- plot_line_rates(
    results[["unemployment"]],
    xcol = "DATA_YEAR",
    legend = "region's unemployment rate"
)

plot(p_unemp)
```

## Economic Opportunity

### Educational Attainment

```{r}
#| label: education_data
rs_ed_attain <- lapply(pumslist_p, psrc_pums_count, group_vars = "ED_ATTAIN", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
results[["education"]] <- rs_ed_attain[ED_ATTAIN != "Total"]
```

```{r}
#| label: education_plot
# p_ed_attain <- plot_rates(
#     filter(rs_ed_attain, ED_ATTAIN != "Total"),
#     group_vars = c("DATA_YEAR", "ED_ATTAIN"),
#     rate_var = "share",
#     whiskers = TRUE, # error bars
#     xlabi = list(size = 2, angle = 90, hjust = 0)
# )

p_ed_attain <- plot_stackbar_rates(
    results[["education"]],
    xcol = "DATA_YEAR",
    ycol = "share",
    fill = "ED_ATTAIN"
)

plot(p_ed_attain)

```

## Quality of Life

### Homeownership

```{r}
#| label: homeownership_data
rs_own_rent <- lapply(pumslist_h, psrc_pums_count, group_vars = "OWN_RENT", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
results[["homeownership"]] <- rs_own_rent[OWN_RENT == "Owned"]
```

```{r}
#| label: homeownership_plot
# p_own_rent <- plot_rates(
#     filter(rs_own_rent, OWN_RENT == "Owned"),
#     group_vars = "DATA_YEAR",
#     rate_var = "share",
#     whiskers = TRUE # error bars
# )

p_own_rent <- plot_line_rates(
    results[["homeownership"]],
    xcol = "DATA_YEAR",
    ycol = "share",
    legend = "region's homeownership rates"
)

plot(p_own_rent)
```

### Renter cost burden

```{r}
#| label: renter_cost_data
rs_rent_burden <- lapply(pumslist_h, psrc_pums_count, group_vars = "rent_burden", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
results[["renter_cost_burden"]] <- rs_rent_burden[
    rent_burden %in% c("Greater than 50 percent", "Between 30 and 50 percent")]
```


```{r}
#| label: renter_cost_plot
# p_rent_burden <- plot_rates(
#     filter(rs_rent_burden, 
#            rent_burden %in% c("Greater than 50 percent",
#                               "Between 30 and 50 percent")),
#     group_vars = c("DATA_YEAR", "rent_burden"),
#     rate_var = "share",
#     whiskers = TRUE, # error bars
#     xlabi = list(size = 2, angle = 90, hjust = 0)
# )

p_rent_burden <- plot_stackbar_rates(
    results[["renter_cost_burden"]],
    xcol = "DATA_YEAR",
    ycol = "share",
    fill = "rent_burden"
)


plot(p_rent_burden)
```

### Lacking Health Insurance

```{r}
#| label: insurance_data
rs_health_insurance <- lapply(pumslist_p, psrc_pums_count, group_vars = "health_insurance", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
results[["no_health_insurance"]] <- rs_health_insurance[
    health_insurance == "No health insurance"]
```

```{r}
#| label: insurance_plot
# p_no_health_insurance <- plot_rates(
#     filter(rs_health_insurance, health_insurance == "No health insurance"),
#     group_vars = "DATA_YEAR",
#     rate_var = "share",
#     whiskers = TRUE # error bars
# )

p_no_health_insurance <- plot_line_rates(
    results[["no_health_insurance"]],
    xcol = "DATA_YEAR",
    ycol = "share",
    legend = "region's lacking insurance rate"
)

plot(p_no_health_insurance)
```

### Lacking Internet Access

```{r}
#| label: internet_data
rs_internet <- NULL
for (i in 1:length(pumslist_h)){
    if(! "internet" %in% colnames(pumslist_h[[i]])) next
    rs_internet[[i]] <- psrc_pums_count(pumslist_h[[i]], 
                                        group_vars = "internet", 
                                        incl_na=FALSE)
}
rs_internet <- data.table::rbindlist(rs_internet, fill=TRUE)
results[["no_internet"]] <- rs_internet[
    DATA_YEAR > 2016 & internet == "Without internet access"]
```

```{r}
#| label: internet_plot
p_internet <- plot_rates(
    results[["no_internet"]],
    group_vars = "DATA_YEAR",
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_internet)
```

## Business Climate

```{r}

```

## Global Competitiveness

```{r}

```

```{r}
# save data
if(save_data)
    write_xlsx(results, path = paste0("annual_comparison_data_", Sys.Date(), ".xlsx"))

```
