---
title: "Annual Trend Data W/O Race & Ethnicity Details"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
library(dplyr)
library(data.table)

# Note: one has to set the data path in get_pums_res_equity.R
lapply(c("get_pums_res_equity.R",
         "plot_rates.R", "plot_counts.R"), source)

dyears <- c(2017:2019, 2021:2024)

pumslist_h <- list()
pumslist_h <- lapply(dyears, get_pums_res_equity_h, span = 1)
pumslist_h <- lapply(pumslist_h, prep_internet_data) %>%
              lapply(prep_renter_burden_data)

pumslist_p <- list()
pumslist_p <- lapply(dyears, get_pums_res_equity_p, span = 1)
pumslist_p <- lapply(pumslist_p, prep_wage_data) %>%
              lapply(prep_labor_force_data) %>%
              lapply(prep_health_insurance_data) %>%
              lapply(prep_poverty_data)
```

## Population Indicators

### Population

```{r}
rs_pop <- lapply(pumslist_p, psrc_pums_count) %>% data.table::rbindlist(fill=TRUE)
```

```{r}
p_pop <- plot_counts(rs_pop, "DATA_YEAR", "count", moe = "count_moe")
plot(p_pop)
```

### Median household income

```{r}
rs_hhincome <- lapply(pumslist_h, psrc_pums_median, stat_var = "HINCP") %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_hhincome <- plot_counts(rs_hhincome, "DATA_YEAR", "HINCP_median", moe = "HINCP_median_moe")
plot(p_hhincome)
```

## Poverty Indicators

### Poverty rates

```{r}
rs_poverty <- lapply(pumslist_p, psrc_pums_count, group_vars = "poverty", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_poverty <- plot_rates(
    filter(rs_poverty, poverty == "Yes"),
    group_vars = "DATA_YEAR",
    whiskers = TRUE # error bars 
)

plot(p_poverty)
```

### Low income rates

```{r}
rs_lowinc <- lapply(pumslist_p, psrc_pums_count, group_vars = "low_income", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_lowinc <- plot_rates(
    filter(rs_lowinc, low_income == "Yes"),
    group_vars = "DATA_YEAR",
    whiskers = TRUE # error bars 
)

plot(p_lowinc)
```

## Employment

### Labor force participation (16 and over)

```{r}
rs_lfp <- lapply(pumslist_p, psrc_pums_count, group_vars = "labor_force_status", incl_na=FALSE) %>%
   data.table::rbindlist(fill=TRUE)
```

```{r}
p_lfp <- plot_rates(
    filter(rs_lfp, labor_force_status == "In labor force"),
    group_vars = "DATA_YEAR",
    whiskers = TRUE # error bars 
)

plot(p_lfp)
```

### Unemployment

```{r}
rs_unemp <- lapply(pumslist_p, psrc_pums_count, group_vars = "employment_status", incl_na=FALSE) %>%
   data.table::rbindlist(fill=TRUE)
```

```{r}
p_unemp <- plot_rates(
    filter(rs_unemp, employment_status == "Unemployed"),
    group_vars = "DATA_YEAR",
    whiskers = TRUE # error bars 
)

plot(p_unemp)
```

## Economic Opportunity

### Educational Attainment

```{r}
rs_ed_attain <- lapply(pumslist_p, psrc_pums_count, group_vars = "ED_ATTAIN", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_ed_attain <- plot_rates(
    filter(rs_ed_attain, ED_ATTAIN != "Total"),
    group_vars = c("DATA_YEAR", "ED_ATTAIN"),
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_ed_attain)

```

## Quality of Life

### Homeownership

```{r}
rs_own_rent <- lapply(pumslist_h, psrc_pums_count, group_vars = "OWN_RENT", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_own_rent <- plot_rates(
    filter(rs_own_rent, OWN_RENT == "Owned"),
    group_vars = "DATA_YEAR",
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_own_rent)
```

### Renter cost burden

```{r}
rs_rent_burden <- lapply(pumslist_h, psrc_pums_count, group_vars = "rent_burden", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)

p_rent_burden <- plot_rates(
    filter(rs_rent_burden, 
           rent_burden %in% c("Greater than 50 percent",
                              "Between 30 and 50 percent")),
    group_vars = c("DATA_YEAR", "rent_burden"),
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_rent_burden)
```

### Lacking Health Insurance

```{r}
rs_health_insurance <- lapply(pumslist_p, psrc_pums_count, group_vars = "health_insurance", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_no_health_insurance <- plot_rates(
    filter(rs_health_insurance, health_insurance == "No health insurance"),
    group_vars = "DATA_YEAR",
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_no_health_insurance)
```

### Lacking Internet Access

```{r}
rs_internet <- lapply(pumslist_h, psrc_pums_count, group_vars = "internet", incl_na=FALSE) %>% 
    data.table::rbindlist(fill=TRUE)
```

```{r}
p_internet <- plot_rates(
    filter(rs_internet, DATA_YEAR > 2016 & internet == "Without internet access"),
    group_vars = "DATA_YEAR",
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_internet)
```

## Business Climate

```{r}

```

## Global Competitiveness

```{r}

```
