---
title: "equity_metrics"
format: html
editor: visual
---

```{r setup, echo=TRUE, output=FALSE}
library(magrittr)
library(dplyr)
library(data.table)

save_data <- TRUE

script_dir <- if (file.exists("get_pums_res_equity.R")) "." else "equity_metrics"
lapply(file.path(script_dir, c("get_pums_res_equity.R", "plot_rates.R", "plot_counts.R", "render_table.R")),
    source, echo = FALSE)

```

```{r data-retrieval}
dyear <- 2024

pumsdata_h <- get_pums_res_equity_h(dyear = dyear)
pumsdata_h <- pumsdata_h %>%
              prep_internet_data() %>%
              prep_renter_burden_data()

pumsdata_p <- get_pums_res_equity_p(dyear = dyear)
pumsdata_p <- pumsdata_p %>%
              prep_wage_data() %>%
              prep_labor_force_data() %>%
              prep_health_insurance_data() %>%
              prep_poverty_data()

```

## Background Metrics

### Population by race/ethnicity

```{r pop}

rs_pop <- psrc_pums_count(pumsdata_p, group_vars = c("PRACE"))
tbl_pop <- filter(rs_pop, PRACE != "Total")

p_pop <- plot_rates(
  tbl_pop,
  group_vars = "PRACE",
  whiskers = TRUE # error bars
)

plot(p_pop)
```

```{r pop-table}
render_data_table(tbl_pop, caption = "Population by Race/Ethnicity")
```

### Median household income by race/ethnicity

```{r median-household-income}

rs_hhincome <- psrc_pums_median(pumsdata_h, stat_var = "HINCP", group_vars = c("HRACE"))
tbl_hhincome <- filter(rs_hhincome, HRACE != "Total")

p_hhincome <- plot_counts(
  tbl_hhincome,
    xcol = "HRACE",
    ycol = "HINCP_median"
)

plot(p_hhincome)
```

```{r median-household-income-table}
render_data_table(tbl_hhincome, caption = "Median Household Income by Race/Ethnicity")
```

### Poverty rates by race/ethnicity

```{r poverty-rates}

rs_poverty <- psrc_pums_count(pumsdata_p, group_vars = c("PRACE", "poverty"), incl_na=FALSE)
tbl_poverty <- filter(rs_poverty, poverty == "Yes")

p_poverty <- plot_rates(
    tbl_poverty,
    group_vars = "PRACE",
    whiskers = TRUE # error bars
)

plot(p_poverty)
```

```{r poverty-rates-table}
render_data_table(tbl_poverty, caption = "Poverty Rates by Race/Ethnicity")
```

### Low income rates by race/ethnicity

```{r low-income-rates}

rs_lowinc <- psrc_pums_count(pumsdata_p, group_vars = c("PRACE", "low_income"), incl_na=FALSE)
tbl_lowinc <- filter(rs_lowinc, low_income == "Yes")

p_lowinc <- plot_rates(
    tbl_lowinc,
    group_vars = "PRACE",
    whiskers = TRUE # error bars
)

plot(p_lowinc)
```

```{r low-income-rates-table}
render_data_table(tbl_lowinc, caption = "Low Income Rates by Race/Ethnicity")
```

### Labor force participation (16-64) by gender and race/ethnicity

```{r unemployment-labor-force-participation}

rs_lfp <- psrc_pums_count(pumsdata_p,
    group_vars = c("PRACE_x_SEX", "labor_force_status"), 
    incl_na=FALSE) %>%
   dplyr::filter(PRACE_x_SEX != "Total") %>%
   tidyr::separate(PRACE_x_SEX, into = c("PRACE", "SEX"), sep = ";")
tbl_lfp <- filter(rs_lfp, labor_force_status == "In labor force")

p_lfp <- plot_rates(
    tbl_lfp,
    group_vars = c("PRACE", "SEX"),
    whiskers = TRUE # error bars
)

plot(p_lfp)
```

```{r labor-force-participation-table}
render_data_table(tbl_lfp, caption = "Labor Force Participation by Gender and Race/Ethnicity")
```

### Unemployment (16+) by gender and race/ethnicity

```{r unemployment-rates}

rs_unemp <- psrc_pums_count(pumsdata_p,
    group_vars = c("PRACE_x_SEX", "employment_status"), 
    incl_na=FALSE) %>%
   dplyr::filter(PRACE_x_SEX != "Total") %>%
   tidyr::separate(PRACE_x_SEX, into = c("PRACE", "SEX"), sep = ";")
tbl_unemp <- filter(rs_unemp, employment_status == "Unemployed")

p_unemp <- plot_rates(
    tbl_unemp,
    group_vars = c("PRACE", "SEX"),
    whiskers = TRUE # error bars
)

plot(p_unemp)
```

```{r unemployment-rates-table}
render_data_table(tbl_unemp, caption = "Unemployment Rates by Gender and Race/Ethnicity")
```

### Economic Opportunity

### Educational Attainment by race/ethnicity

```{r educational-attainment}

rs_ed_attain <- psrc_pums_count(pumsdata_p, 
    group_vars = c("PRACE", "ED_ATTAIN"), 
    incl_na=FALSE)
tbl_ed_attain <- filter(rs_ed_attain, ED_ATTAIN != "Total")

p_ed_attain <- plot_rates(
    tbl_ed_attain,
    group_vars = c("PRACE", "ED_ATTAIN"),
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_ed_attain)
```

```{r educational-attainment-table}
render_data_table(tbl_ed_attain, caption = "Educational Attainment by Race/Ethnicity")
```

### Kindergarten readiness by race/ethnicity

```{r kindergarten-readiness}
#-- direct sourcing; uses RSocrata `remotes::install_github("Chicago/RSocrata")` 
source("https://raw.githubusercontent.com/psrc/equity-tracker/refs/heads/main/data_development/get_ospi_equity_indicators.R")

ospi_kready <- get_k_readiness("https://data.wa.gov/resource/veeh-ue57.json") # 2025-26 school year
colnames(ospi_kready) <- tolower(colnames(ospi_kready))
ospi_kready <- ospi_kready %>% filter(studentgrouptype == "FederalRaceEthnicity")
rs_kready <- ospi_kready %>% 
  group_by(studentgroup) %>% 
  summarise(numerator=sum(numerator, na.rm=TRUE),
            denominator=sum(denominator, na.rm=TRUE), 
            .groups = "drop") %>% 
  mutate(share=numerator/denominator)

p_kready <- plot_rates(rs_kready,
    group_vars = "studentgroup",
    whiskers = FALSE # no MOE for this metric
)  

plot(p_kready)

```

```{r kindergarten-readiness-table}
render_data_table(rs_kready, caption = "Kindergarten Readiness by Race/Ethnicity")
```

## Quality of Life

### Homeownership by race/ethnicity

```{r homeownership}

rs_own_rent <- psrc_pums_count(pumsdata_h, 
    group_vars = c("HRACE", "OWN_RENT"), 
    incl_na=FALSE)
tbl_own_rent <- filter(rs_own_rent, OWN_RENT == "Own")

p_own_rent <- plot_rates(
    tbl_own_rent,
    group_vars = c("HRACE"),
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_own_rent)
```

```{r homeownership-table}
render_data_table(tbl_own_rent, caption = "Homeownership by Race/Ethnicity")
```

### Renter cost burden by race/ethnicity

```{r renter-cost-burden}
rs_rent_burden <- psrc_pums_count(pumsdata_h, 
    group_vars = c("HRACE", "rent_burden"), 
    incl_na=FALSE)
tbl_rent_burden <- filter(rs_rent_burden, 
           rent_burden %in% c("Greater than 50 percent",
                              "Between 30 and 50 percent"))

p_rent_burden <- plot_rates(
    tbl_rent_burden,
    group_vars = c("HRACE", "rent_burden"), 
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_rent_burden)
```

```{r renter-cost-burden-table}
render_data_table(tbl_rent_burden, caption = "Renter Cost Burden by Race/Ethnicity")
```

### Health insurance coverage by race/ethnicity

```{r health-insurance}
rs_health_insurance <- psrc_pums_count(pumsdata_p, 
    group_vars = c("PRACE", "health_insurance"), 
    incl_na=FALSE)
tbl_health_insurance <- filter(rs_health_insurance, health_insurance == "No health insurance")

p_no_health_insurance <- plot_rates(
    tbl_health_insurance,
    group_vars = c("PRACE"),
    rate_var = "share",
    whiskers = TRUE # error bars
)

plot(p_no_health_insurance)
```

```{r health-insurance-table}
render_data_table(tbl_health_insurance, caption = "Lack of Health Insurance by Race/Ethnicity")
```

### Internet Access by race/ethnicity

```{r internet-access}
rs_internet <- psrc_pums_count(pumsdata_h, group_vars = c("HRACE", "internet"))
tbl_internet <- filter(rs_internet, internet != "Total")

p_internet <- psrcplot::static_bar_chart(
    t = tbl_internet,
    y = "HRACE",
    x = "share",
    fill = "internet",
    title = "Internet Access by Race/Ethnicity",
    alt = "Chart of Internet Access by Race/Ethnicity",
    source = paste("Source: ACS PUMS 2020-2024 5-year microdata",
                    "for King, Kitsap, Pierce and Snohomish counties.",
                    sep = "\n"),
    color="pgnobgy_5", pos = "stack")

plot(p_internet)
```

```{r internet-access-table}
render_data_table(tbl_internet, caption = "Internet Access by Race/Ethnicity")
```

## Business Climate

```{r business-climate-plots}

```

## Global Competitiveness

```{r global-competitiveness-plots}

```

##Write tables to file

```{r save-results-to-file, echo=TRUE}
if(save_data){
    library(writexl)
    results <- list() # for collecting all results to store as spreadsheet
    results$pop <- rs_pop
    results$hh_income <- rs_hhincome
    results$poverty <- rs_poverty
    results$lowinc <- rs_lowinc
    results$lfp <- rs_lfp
    results$unemp <- rs_unemp
    results$ed_attain <- rs_ed_attain
    results$kready <- rs_kready
    results$own_rent <- rs_own_rent
    results$rent_burden <- rs_rent_burden
    results$health_insurance <- rs_health_insurance
    results$internet <- rs_internet

    write_xlsx(results, path = paste0("equity_metrics", Sys.Date(), ".xlsx"))
}

```